---
title: "DSC520_Final_Proj_Quinn"
author: "Zach Quinn"
date: "5/27/2020"
output: html_document
---

Since this project endeavors to distinguish between the impact of new and legacy content on the Disney Plus platform, a comprehensive statistical and graphical analysis will be conducted to guide the following inquiry: What film content do Disney Plus subscribers prefer more, the so-called Disney classics or Disney Plus Original programming? 

```{r}

# Displays first six rows of classic films data set. 

# This project considers a “Disney Classic Film” to correspond to the 56 officially sanctioned films. A “Disney Plus Original Film” is any content produced within the last five years (2015 – 2020) that is native to the Disney Plus platform. In the course of data preparation, the original data set was split into these two distinct data.frames, one of which, “Disney.Classic.Film” corresponds to the cannon of Walt Disney Classics and “Disney.Plus.Original.Film” corresponds to values that fit the parameters of a Disney Plus Original. The Disney Classic data set was reduced to 52 observations of 7 variables, each hand-selected to exclude films that are not explicitly considered a Disney Classic by the Walt Disney Company.

head(Disney.Classic.Films)

```

```{r}

# Displays first six rows of original films data set. 

# The Disney Plus Original data set contains 43 observations of 7 variables. The variables included in each data set are: Title (factor), MPAA rating (factor), genre (factor), year released (numeric), IMDB votes (numeric), IMDB rating (numeric), and the number of awards (numeric).  

head(Disney.Plus.Original.Films)

```
```{r}

# Displays summary statistics for classic films data set. 

summary(Disney.Classic.Films)

```
```{r}

# Displays summary statistics for original films data set. 

summary(Disney.Plus.Original.Films)
```
```{r}

# Calculates correlation statistics for variable relationships.

cor.classic.films <- cor(Disney.Classic.Films[,c('Year', 'IMDB.Rating', 'IMDB.Votes', 'Awards')])
cor.classic.films

```
```{r}
ggcorrplot(cor.classic.films, hc.order = TRUE, lab = TRUE) + ggtitle('Correlation of Disney Classic Films')

# The correlation matrix demonstrates differing degrees of correlation on the four numeric variables involved in this data set analysis. Overall, the Disney Classics data set demonstrated a higher degree of correlation between the selected variables. The “Correlation of Disney Films” correlation matrix suggests that IMDB votes and IMDB ratings have a strong positive correlation at 0.63. Release year has a slightly positive impact upon the number of IMDB votes recorded for a particular classic Disney title at a correlation level of 0.35. Additional subtle positive correlations exist between Awards and IMDB rating at 0.11. Negative correlations are evident among the awards and release year variables (-0.04) and zero values were recorded for the year and award variables, suggesting a neutral correlation between these variables. The positive correlation of IMDB votes and ratings are illustrated in a scatter plot included in this analysis. 

```
```{r}

# Calculates correlation between IMDB rating and IMDB votes for classic films data set. 

cor.test(IMDB.Rating, IMDB.Votes)
```
```{r}

# The graph demonstrates a positive correlation between the two variables extracted from the IMDB data. Notably, content that receives a higher rating tends to receive more votes. This suggests that viewers are more likely to view established content that has already received hundreds of thousands of votes and is, therefore, tacitly endorsed by fellow viewers. Additionally, the year coloration reveals data trends that contradict popular preconceptions of Disney content, including Disney marketing itself, which lauds such ‘timeless classics.’ Instead, the data suggests that both the highest rated and the most voted-upon content was that which was released from 2000 onward.

ggplot(Disney.Classic.Films, aes(x = IMDB.Rating, y = IMDB.Votes, col = Year)) + geom_point() + ggtitle('Disney Classic Films IMDB Data') + geom_smooth(method = lm, se = FALSE)

```
```{r}
# Constructing a linear regression model enables further exploration of the impact of the variables in the following formula IMDB.Rating ~ IMDB.Votes. 

summary(lm(IMDB.Rating ~ IMDB.Votes, data = Disney.Classic.Films))

#  The statistical summary suggests there is a strong linear relationship between the selected variables. The t-value is large relative to the negative standard of error. Additionally, the fact that the p-value is close to zero suggests that it is plausible to reject the null hypothesis for this formula, meaning that there is, indeed, a positive correlation between the variables of IMDB rating and IMDB votes. 

```
```{r}

ggplot(Disney.Classic.Films, aes(x = Year, y = IMDB.Votes, color = Genre)) + geom_point() + ggtitle('Disney Classics Votes vs. Year')

# The below graph somewhat contradicts this paper’s initial hypothesis, that older ‘classic’ content would be more popular. Instead, while there are significant clusters of data between the years of 1950 and 1975, the most popular content is mostly concentrated between the year 2000 and the present. Additionally, while Disney’s earliest content was primarily situated in the genres of drama and musical. Comparably modern classics have pivoted toward romance and comedy, more widely appealing modes of entertainment. 

```
```{r}
summary(lm(IMDB.Rating ~ IMDB.Votes, data = Disney.Classic.Films))

# A summary of the regression model Votes ~ Year is not as statistically significant as the model IMDB.Rating ~ Votes. However, its p-values, both at 0.01 suggest that a positive correlation exists among the variables and it is still possible to reject the null hypothesis. 

```

```{r}
# While the correlation among the variables in the Disney Classic Films data frame was conducive to an analysis of subscriber trends, there was a lesser degree of correlation among the variables in the Disney Plus Originals correlation matrix.

cor.original.films <- cor(Disney.Plus.Original.Films[,c('OG.Year', 'OG.IMDB.Rating', 'OG.IMDB.Votes', 'OG.Awards')])
cor.original.films
```
```{r}
ggcorrplot(cor.original.films, hc.order = TRUE, lab = TRUE) + ggtitle('Disney Plus Originals Correlation')

# For instance, while the Disney Classic Films data set’s highest degree of correlation was 0.63, the highest degree of correlation among the variables in the Disney Plus Originals correlation matrix is 0.32. In this case, the correlation in question was among the variables of IMDB votes and original content awards. The second highest correlation trend was observed between IMDB votes and IMDB ratings for original films hosted on the Disney Plus platform. The least significant positive correlation exists among IMDB votes and year at 0.01. Negative correlations were reported between year and IMDB rating (-0.1) as well as IMDB rating and awards (-0.1) and year and awards (-0.13).

```
```{r}
# Calculates correlation between numeric variables awards and IMDB votes. 

cor.test(Disney.Plus.Original.Films.Awards, Disney.Plus.Original.Films.IMDB.Votes)

```
```{r}

summary(lm(Disney.Plus.Original.Films.IMDB.Votes ~ Disney.Plus.Original.Films.Awards))

# Though the p-value is small, it approaches the 0.05 threshold. So, while it is possible to reject the null hypothesis, this is not a considerably significant correlation.

```

```{r}
ggplot(Disney.Plus.Original.Films, aes(x = Disney.Plus.Original.Films.Awards, y = Disney.Plus.Original.Films.IMDB.Votes)) + geom_col() + ggtitle('Disney Plus IMDB Votes vs. Awards')

# The graph for the most significant correlation, IMDB votes vs. Awards is detailed in the histogram below. The above graph exhibits a regressive relationship between the variables, as votes on the content decreases as the number of awards increase. Though the regression appears substantial on the graph, a summary of the linear model Disney.Plus.Original.Films.IMDB.Votes ~ Disney.Plus.Original.Films.Awards reveals that the relationship is only somewhat significant. 

```
```{r}
ggplot(Disney.Plus.Original.Films, aes(x = Disney.Plus.Original.Films.Genre, y = Disney.Plus.Original.Films.IMDB.Rating)) + geom_col(fill = "#00FFFF") + ggtitle('Disney Plus Originals Genre Preferences') + coord_flip()

# Unlike the Disney Classics data set, the Disney Plus Original films data set genre preferences reveal that the original content that resonates most with subscribers is the documentary genre. The fantasy, sci-fi and comedy genres narrowly follow the documentary category in popularity. This is a significant departure from the classic data set, which exclusively measures fictional content. However, being a categorical variable, this graph only broadly measures genre preference without yielding a significant summary statistic such as a p-value, t-test or r-squared. 

```

Therefore, after analyzing both data sets, the statistically significant correlations hinge on IMDB data, including the total number of votes and rating viewers assigned to each title. Additionally, for the Disney Classics data set, which spans over 60 years, year is a relevant variable when juxtaposed with IMDB votes and colored with genre. Given the limited availability of content and, consequently, raw data, related to the Disney Plus Original data set, the only statistically significant relationship observed was between IMDB votes and awards. 

From this analysis, it is reasonable to conclude that the year content was released and numbers of IMDB votes are the most relevant predictor variables in determining the outcome that a viewer is likely to prefer Disney classic films over Disney Plus original films. Based solely upon the maximum values in both data sets’ IMDB votes column, Disney classic films had a higher maximum value at 863,619 to the Disney Plus’ data set’s maximum value of 694,527. However, in both data sets, the newest content received the largest number of IMDB votes and highest IMDB ratings. While Disney Plus’ original content boasts promising offerings, in its current state, it simply lacks the scope and prestige of the robust Disney classic films catalogue. Based on the above analysis, the correlations between IMDB data and classic Disney content is more statistically significant than that exhibited by the Disney Plus original content data set. Therefore, the data suggests that Disney Plus subscribers marginally prefer content classified as classic than that classified as Disney Plus originals. 

